---
title: "Data Cleaning"
author: "Katherine Nishimura"
date: "July 12, 2016"
output:
  pdf_document: default
  html_document: default
---

### Notes from Module 1, Day 2, afternoon session 1
- 

### A. "Real Property Taxes" Dataset 

Download the "real property taxes" dataset from the website (via OpenBaltimore), 
he data is located here (note you don\'t need to unzip it to read it into R):
http://sisbid.github.io/Module1/data/real_property_tax.csv.gz

1. Read the Property Tax data into R and call it the variable `tax`
- read.table to .gz documents (most, but not all gzipped .csv files)
- read_csv(file.gz)

##### install.packages("readr")

```{r chunk1, echo=TRUE}
library(readr)
data_url <- "http://sisbid.github.io/Module1/data/real_property_tax.csv.gz"
data <- read_csv(data_url)
head(data)
dim(data)
names(data)
```

2. How many addresses pay property taxes? 

```{r, chunk2, echo=TRUE}
#convert cityTax and stateTax variables to numeric
class(data$cityTax) #character
#replace "$" with ""
data$cityTax <- gsub(pattern = "$", replacement="", data$cityTax, fixed=TRUE) 
data$stateTax <- gsub(pattern = "$", replacement="", data$stateTax, fixed=TRUE)
# convert to numeric
data$cityTax <- as.numeric(data$cityTax)
data$stateTax <- as.numeric(data$stateTax)
# check to see how many addresses pay property taxes
summary(data$cityTax)
summary(data$stateTax)
data <- data[order(data$cityTax, decreasing=FALSE), ]
dim(data)
no_property_tax <- subset(data, data$cityTax < 1 )
dim(no_property_tax) #72 rows where cityTax < 1
rm(no_property_tax)
```

3. What is the total city and state tax paid?

```{r, chunk3, echo=TRUE}
#total city and state tax paid
head(data)
# generate column sums for columns in "data" that are numeric, removing NA values
colSums(Filter(is.numeric, data), na.rm=TRUE) 
```


4. How many observations/properties are in each residence code?

```{r, chunk4, echo=TRUE}
# observations per residence code (resCode)
table(data$resCode)
```


5. What is the 75th percentile of city and state tax paid by residence code?

```{r, chunk5, echo=TRUE}
# 75th percentile of city and state tax paid by residence code
summary(data$cityTax, na.rm=TRUE)
summary(data$stateTax, na.rm=TRUE)
```


6. Subset the data to only retain those houses that are principal residences and describe the distribution of property taxes on these residences.

```{r, chunk6, echo=TRUE}
# data subset with only principal residences
table(data$resCode)
data_principal <- subset(data, data$resCode == "PRINCIPAL RESIDENCE")
dim(data)
dim(data_principal)
# property taxes of these residences
summary(data_principal$cityTax, na.rm=TRUE)
summary(data_principal$stateTax, na.rm=TRUE)
```

7. Convert the 'lotSize' variable to a numeric square feet variable. Tips:
* Look at the data 
* Assume hyphens represent inches within square foot meassuremnts 
* Assume decimals within acreage measurements
* 1 acre = 43560 square feet
* Look at the data 

```{r, chunk7, echo=TRUE}
# lotSize variable
head(data)
data[100:120, c("lotSize")]

```



### B. "Baltimore Salary 2015" Dataset 

Download the "Salary 2015" Dataset from the website (via OpenBaltimore), which is located here:
http://sisbid.github.io/Module1/data//Baltimore_City_Employee_Salaries_FY2015.csv

8. Make an object called health.sal using the salaries data set,	with only agencies of those with "fire" (or any forms), if any, in the name

##### install.packages("readr")
##### install.packages("stringr")
##### .


```{r, chunk8, echo=TRUE}
library(readr)
data_url <- "http://sisbid.github.io/Module1/data//Baltimore_City_Employee_Salaries_FY2015.csv"
data <- read_csv(data_url)
head(data)
dim(data)
names(data)
# subset rows where any word in "Agency" variable is "fire"
data_subset <- data[grep("Fire", data$Agency), ] 
head(data_subset)
rm(data_subset)
```


9. Make a data set called trans which contains only agencies that contain "TRANS".

```{r, chunk9, echo=TRUE}
# subset rows where any word in "Agency" variable is "TRANS"
data_subset <- data[grep("TRANS", data$Agency), ] 
dim(data)
dim(data_subset)
head(data_subset)
rm(data_subset)
```


10. What is/are the profession(s) of people who have "abra" in their name for Baltimore's Salaries?

install.packages("stringr")

```{r, chunk10, echo=TRUE}
library(stringr)
# people with "abra" in their name
names(data)
data$NAMES <- data[, 1]
data_subset <- data[grep("Aaron", data$NAMES), ] 
dim(data)
dim(data_subset)
head(data_subset)
rm(data_subset)
```


11. What is the distribution of annual salaries look like? What is the IQR?

```{r, chunk11a, echo=TRUE}
names(data)
head(data)
# replace "$" with ""
data$AnnualSalary <- gsub(pattern = "$", replacement="", data$AnnualSalary, fixed=TRUE)
# convert to numeric
data$AnnualSalary <- as.numeric(data$AnnualSalary)
# summarize
summary(data$AnnualSalary, na.rm=TRUE)
```

- histogram 
```{r, chunk11b, echo=TRUE, fig.height=5}
hist(data$AnnualSalary)
```


12. Convert HireDate to the `Date` class - plot Annual Salary vs Hire Date

```{r, chunk12a, echo=TRUE}
names(data)
head(data)
# convert "HireDate" to a date class variable
# %m = numerical month 1-12; %d = numerical day 1-31; %Y = 4 digit year
data$HireDate <- as.Date(data$HireDate, "%m/%d/%Y")
class(data$HireDate)
```

13. Plot annual salary versus hire date.	Hint: first convert to numeric and date respectively
- plot annual salary vs hire date


```{r, chunk12b, echo=TRUE, fig.height=5}
names(data)
head(data)
plot(data$HireDate, data$AnnualSalary)
```



14. Create a smaller dataset that only includes the Police Department, Fire Department and Sheriff's Office. How many employees are in this new dataset?

```{r, chunk14, echo=TRUE}
names(data)
head(data)
data_subset <- data[grep("Police|Fire|Sheriff", data$Agency), ] #subset if "Agency" includes Police, Fire, or Sherrif
dim(data)
dim(data_subset)
head(data_subset)
rm(data_subset)
```




### Include the session tools
```{r 10, echo=TRUE}
devtools::session_info()
```



